---
title: "CodeBook"
output: html_document
---

Objective 1 of this R script is to merge the training and test sets to create one data set

Objective 2 of this R script is to extract columns containing mean and standard deviation for each measurement

Objective 3 of this script is to create variables called ActivityLabel and ActivityName that label all observations with the corresponding activity labels and names respectively

Objective 4 of this script is to make a tidy data set with the average of each variable for each activity and each subject.

```{r}
features <- read.table (file.choose ())
subject_test <- read.table (file.choose ())
subject_train <- read.table (file.choose ())
x_train <- read.table (file.choose ())
Y_train<- read.table(file.choose())

X_test <- read.table (file.choose ())
Y_test <- read.table (file.choose ())

body_acc_X_train <- read.table (file.choose ())
body_acc_Y_train <- read.table (file.choose ())
body_acc_Z_train <- read.table (file.choose ())

body_gyro_X_train <- read.table (file.choose ())
body_gyro_Y_train <- read.table (file.choose ())
body_gyro_Z_train <- read.table (file.choose ())

total_acc_X_train <- read.table (file.choose ())
total_acc_Y_train <- read.table (file.choose ())
total_acc_Z_train <- read.table (file.choose ())

body_acc_X_test <- read.table (file.choose ())
body_acc_Y_test <- read.table (file.choose ())
body_acc_Z_test <- read.table (file.choose ())

body_gyro_X_test <- read.table (file.choose ())
body_gyro_Y_test <- read.table (file.choose ())
body_gyro_Z_test <- read.table (file.choose ())

total_acc_X_test <- read.table (file.choose ())
total_acc_Y_test <- read.table (file.choose ())
total_acc_Z_test <- read.table (file.choose ())
```

Above mentioned code will prompt the window to load each file step by step, I have done this so as it can be run on any system, otherwise, you can specify the file paths on your system.

Next step would be to merge the data frames to make a single data set.

I am going to use merge () function for merging the x_test and X_train datasets

``` {r}
x_train_test_merged <- merge (X_test, x_train, all = TRUE)
```

Now we can name the columns using the features data set with the help of colnames function

``` {r}
colnames (x_train_test_merged) <- features$V2
```

We will use bind_rows () for merging Y_test and Y_train as both the data frames have numeric type data.

``` {r}
Y_train_test_merged <- bind_rows (Y_test, Y_train)
```

Now merging both these variables to get one merged data frame.

``` {r}
consolidated_x_y <- data.frame (x_train_test_merged,Y_train_test_merged)

subject_total <- bind_rows (subject_test, subject_train)

consolidated_x_y <- data.frame (subject_total, consolidated_x_y)

colnames (consolidated_x_y) [1] <- "Subjects"
```

With above mentioned code we have achieved the objective 1 of the project.

Following code is to achieve objective 2

I have used grep function to extract the text with the strings "mean" and "std"

``` {r}
std_consolidated <- grep ("std", colnames(consolidated_x_y))

mean_consolidated <- grep ("mean", colnames (consolidated_x_y))

consolidated_mean_std <- data.frame (consolidated_x_y [,std_consolidated], consolidated_x_y [, mean_consolidated])
```
Next step would be to define ActivityLabels and ActivityNames so as to label and name the observations

``` {r}
ActivityLAbel <- Activity_Label$V1 [consolidated_x_y$V1]

ActivityName <- Activity_Label$V2 [consolidated_x_y$V1]
```

Functions mentioned above will generate two vectors, first one a numeric vector containing all the labels and second one a factor vector with all the names of the activities

Now I will present the piece of code for the last objective of the project.

``` {r}
install.packages ("tidyr")
```

Now we will use aggregate function for finding out the average of each variable for each activity and each subject

``` {r}
df1 <- aggregate (consolidated_x_y, by= list (Subjects, V1.1), FUN = mean)

gathered_final <- df1  %>% tidyr::gather(Subjects, V1.1, 4:564)
```

WE will now rename the columns for gathered_final variable

```{r}
colnames (gathered.final)[6] <- "Mean_Observations"

colnames (gathered_final)[5] <- "Variables"
```

We are now done with all the objectives of producing this script. we will export it to a data table for submission.

``` {r}
write.table (gathered_final, "C:/Users/Kshitiz Khatri/Desktop/Springboard/getdata-projectfiles-UCI HAR Dataset/tidy.txt", sep="\t")
```

# Thank You


